---
title: "Visualization"
subtitle: "Part II"
output: html_document
date: "2025-09-30"
---

```{r setup}
library(tidyverse)
library(patchwork)
```


# Load the weather data

```{r}
library(p8105.datasets)
data("weather_df")
```


## Remember this plot...?

```{r}
weather_df %>%
  ggplot(aes(x=tmin,y=tmax, color=name)) +
  geom_point(alpha = .5)
```

## Lables

```{r}
weather_df %>%
  ggplot(aes(x=tmin,y=tmax, color=name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from rnoaa package; temperature in 2021."
  )
```

## Scales

Start with the same plot

```{r}
weather_df %>%
  ggplot(aes(x=tmin,y=tmax, color=name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from rnoaa package; temperature in 2021."
  )+
  scale_x_continuous(
    breaks = c(-15, 0, 15),
    labels = c("-15C", "0", "15")
  )+
  scale_y_continuous(
    position = "right"
  )
```

(可以在scale_y_continuous中使用`trans = "log"`将函数拟合进行转换)


Look at color scales

```{r}
weather_df %>%
  ggplot(aes(x=tmin,y=tmax, color=name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from rnoaa package; temperature in 2021."
  )+
  scale_color_hue(
    name = "Location",
    h=c(100,300))
```
(`name = "Location"`:修改颜色图例的标题;
  `h=c(100,300)`:控制颜色的色相范围)

```{r}
weather_df %>%
  ggplot(aes(x=tmin,y=tmax, color=name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from rnoaa package; temperature in 2021."
  )+
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE)
```
(这里`color = name`，name是一个离散型变量，因此要加`discrete = TRUE`)




## Themes

Shift the legend:

```{r}
weather_df %>%
  ggplot(aes(x=tmin,y=tmax, color=name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from rnoaa package; temperature in 2021."
  )+
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE)+
  theme(legend.position = "bottom")
```

Change the overall theme.
```{r}
weather_df %>%
  ggplot(aes(x=tmin,y=tmax, color=name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from rnoaa package; temperature in 2021."
  )+
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE)+
  theme_minimal()
```

 (You can try `theme_classic()`,`ggthemes::theme_excel()`...)


### Setting options

```{r, eval=FALSE}
 library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width= "90%"
)

theme_set(theme_minimal()+theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_color_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Data args in `geom`

```{r}
central_park = 
  weather_df %>%
  filter(name == "CentralPark_NY")


waterhole =
  weather_df %>%
  filter(name == "Waterhole_WA")

ggplot(data = waterhole, aes(x = date, y = tmax, color =name))+
  geom_point()+
  geom_line(data = central_park)
```

  
